Who Needs Privacy?

### 2019-02-27 manuscript/manuscript.RmD
- Integrated last edits by MM

abstract          : 'This study analyzes how personality traits relate to need for privacy. We focus on three dimensions: (a) need for privacy from government surveillance, (b) need for privacy from other individuals, and (c) need for anonymity. Using an online questionnaire with 273 student respondents, several significant relationships were found. For example, results showed that less sociable people desired considerably more privacy on all three dimensions. Somewhat more controversially, people who reported less integrity also desired more anonymity. While anxious people desired slightly less privacy from government surveillance, risk avoidant people desired considerably more privacy from other people. Traditionalism did not relate to need for privacy. Together, the results have several implications. For example, they shed light on the "nothing-to-hide argument": A person who desires more anonymity might indeed be of less integrity. The most important predictor of need for privacy, however, was sociability. In conclusion, this study is the first to show that if a person desires more privacy, it is most plausible that he or she is less sociable.'
keywords          : "Privacy, need for privacy, personality, anonymity, integrity, SEM"

@@

Privacy as a theoretical concept is both complicated and contested [@Nissenbaum.2010, p. 71]. Hence, we begin by outlining our own understanding of privacy. First and foremost, privacy captures the extent of voluntary withdrawal from others [@Westin.1967]. Several models suggest that privacy is multi-dimensional. For example, in a theory-driven treatise @Burgoon.1982 argued that privacy has four dimensions: informational, social, psychological, and physical privacy. @Pedersen.1979, by contrast, argues that on the basis of an empirical factor analysis (initially starting with 94 items) privacy exists on six dimensions: reserve, isolation, solitude, intimacy with friends, intimacy with family, and anonymity. In addition, @Schwartz.1968 differentiated between horizontal and vertical privacy; whereas horizontal privacy captures withdrawal from peers, vertical privacy addresses withdrawal from superiors or institutions (e.g., government agencies). The need for privacy, then, captures "[a]n individual’s need to selectively control the
access of others to the individual = self with the aim of achieving a desired level of physical or psychological privacy [...]") [@Trepte.2017d, p. 1]. Finally, one can also distinguish between the objective privacy context (e.g., how many users can actually read a post on an SNS?), subjective privacy perceptions (e.g., how many users people _think_ can read a post on an SNS?), and resulting privacy behaviors (e.g., how much do people self-disclose?) [@Dienlin.2014].

@@

In order to guide the selection of personality dimensions that might explain need for privacy best, it is useful to determine why people actually feel they need privacy. According to @Trepte.2017d, the need for privacy is a secondary need---it is not an end in itself, but rather a way to satisfy other more fundamental needs. Specifically, @Westin.1967 defined four ultimate purposes of privacy: (1) self-development (i.e., the integration of experiences into meaningful patterns), (2) autonomy (i.e., the desire to avoid being manipulated and dominated), (3) emotional release (i.e., the release of tension from social role demands), and (4) protected communication (i.e., the ability to foster intimate relationships). Not least, privacy facilitates self-disclosure [@Dienlin.2014], which is vital for attaining social support, initiating relationships, and getting close to other people [@Omarzu.2000].

@@

# Methods

@@

Summarizing across the results, the predictors explained `r inspect(results_threefactors_fit, 'r2')["pri_nee_gov"] * 100 %>% round(0)`% of the variance in need for privacy from the government, `r inspect(results_threefactors_fit, 'r2')["pri_nee_int"] * 100 %>% round(0)`% of the variance in need for privacy from other people, and `r inspect(results_threefactors_fit, 'r2')["pri_nee_ano"] * 100 %>% round(0)`% of the variance in need for anonymity. All predictors were also above our predefined SESOI of |$\upbeta$| = .10, supporting their theoretical relevance. In addition, note that several nonsignificant effects were actually larger than the predefined SESOI (e.g., the potential effect of risk avoidance on the need for privacy from the government). Hence, it seems plausible that a larger study, which yields smaller standard errors, would find additional effects.

As stated above, because all predictors were included on the basis of specific a priori rationales we did not control for alpha error inflation. To better assess whether the results could also be explained by chance alone, let us briefly reconsider that if the null hypotheses had been true, given the predefined alpha level of 5 % and the `r nrow(t_results_threefactors) %>% as.english()` statistical tests that were run, one would have expected to find `r round(nrow(t_results_threefactors) * .05, 2)` statistically significant results by chance alone. In this study we found `r round(nrow(filter(coeffs_tab(results_threefactors_fit), p < .05)), 0) %>% english()` statistically significant results, which is substantially higher. Similarly, the average _p_-value _p_ = `r coeffs_tab(results_threefactors_fit)$p %>% mean %>% my_round("std")` was substantially smaller than the average _p_-value of _p_ = .50 that one would have expected to find if the null hypotheses had been true.

@@

Next, and more controversial, our results also showed that the desire for anonymity is related substantially with measures of integrity. People who self-reported lower integrity desired more anonymity. Respondents who said, for example, that they would feel tempted to take things that do not belong to them were also more likely to avoid situations in which they were identifiable. This finding hence supports the logic behind the “nothing to hide” argument. It also follows Altman’s privacy regulation theory [-@Altman.1976], which states that if exposure of information is risky it is likely that people will use more mechanisms to strengthen their social boundaries and increase their desired level of privacy.	Importantly, however, integrity was not related to need for privacy from other people or to need for privacy from government surveillance, which hence limits the generalisability of the nothing-to-hide argument. In addition, note that the responses to the items measuring need for anonymity were much lower compared to the other two dimensions of need for privacy (i.e., _M_~anonymity~ = `r t_psyc["Anonymity", "m"]`, _M_~government~ = `r t_psyc["Government", "m"]`, _M_~interpersonal~ = `r t_psyc["Interpersonal", "m"]`). Hence, the relationship between integrity and need for privacy might show only in very specific domains, for example, when it comes to avoiding places with video surveillance. In conclusion, the relation between integrity and need for privacy seems to be limited to certain aspects of need for privacy only. It should thus not be overly generalized.

@@

Finally, in the SEM no significant relations between need for privacy and traditionality were found. The results imply that independent from being more or less open to change, people desire the same level of privacy. Taken further, this finding might illustrate that people who are more open to change, who are more likely to use novel devices, media, and services, or who might consider themselves as early adopters, do not automatically need less privacy.

@@

Another limitation of the study is that despite using mostly well-established scales, confirmatory factor analyses revealed that most measures did not show satisfactory factorial validity, which resembles the current finding and criticism by @Hussey.2018. In order to achieve sufficient factorial validity we therefore needed to adapt the scales (for example by deleting items). However, this introduces problems of overfitting, which in turn potentially impedes the generalisability of the results [see, e.g., @Yarkoni.2017]. By using bifactor models we explicitly tried to retain the largest possible number of items, which should render our results more robust. In addition, also when looking at the items individually (see the OSM) one can see that the results are not overly dependent on inclusion of specific items.

Another methodological improvement for future studies concerns the sample, which was comparatively small, leading to a reduced power to detect also small effects. Nonetheless, we are still confident that the results should be reliable because we deliberately selected specific predictors that should exhibit at least small to moderate relations with need for privacy (i.e., _r_ = .20). For small to moderate relations, the current study still had a convincing power of `r power_smallmoderate`%. Granted, when aiming to find also small effects (i.e., _r_ = .10) with a probability of 95%, future studies would need to collect data from `r sample_95` respondents [@Cohen.1992]. Finally, generalisability of our sample is limited, as it was predominantly young, female, highly educated, and collected only in one country.

>>>>

abstract          : 'This study analyzes how personality traits relate to need for privacy. We focus on three dimensions: (a) need for privacy from government surveillance, (b) need for privacy from other individuals, and (c) need for anonymity. Using an online questionnaire with 273 student respondents, several significant relationships were found. For example, results showed that less sociable people desired considerably more privacy on all three dimensions. Somewhat more controversially, people who reported less integrity also desired more anonymity. While anxious people desired slightly less privacy from government surveillance, risk avoidant people desired considerably more privacy from other people. Traditionalism did not relate to need for privacy. Together, the results have several implications. For example, they shed light on the "nothing-to-hide argument": A person who desires more anonymity might indeed be of less integrity. The most important predictor of need for privacy, however, was sociability. Indeed, this study is the first to show empirically that if a person desires more privacy, it is most plausible that he or she is less sociable.'
keywords          : "Privacy, need for privacy, personality, anonymity, integrity, SEM"

@@

Privacy as a theoretical concept is both complicated and contested [@Nissenbaum.2010, p. 71]. Hence, we begin by outlining our own understanding of privacy. First and foremost, privacy captures the extent of voluntary withdrawal from others [@Westin.1967]. Several models suggest that privacy is multi-dimensional. For example, in a theory-driven treatise @Burgoon.1982 argued that privacy has four dimensions: informational, social, psychological, and physical privacy. @Pedersen.1979, by contrast, argues that on the basis of an empirical factor analysis (initially starting with 94 items) privacy exists on six dimensions: reserve, isolation, solitude, intimacy with friends, intimacy with family, and anonymity. In addition, @Schwartz.1968 differentiated between horizontal and vertical privacy; whereas horizontal privacy captures withdrawal from peers, vertical privacy addresses withdrawal from superiors or institutions (e.g., government agencies). The need for privacy, then, captures "[a]n individual’s need to selectively control the
access of others to the individual self with the aim of achieving a desired level of physical or psychological privacy [...]") [@Trepte.2017d, p. 1]. Finally, one can also distinguish between the objective privacy context (e.g., how many users can actually read a post on an SNS?), subjective privacy perceptions (e.g., how many users people _think_ can read a post on an SNS?), and resulting privacy behaviors (e.g., how much do people self-disclose?) [@Dienlin.2014].

@@

In order to guide the selection of personality dimensions that might explain need for privacy best, it is useful to determine why people actually feel they need privacy. According to @Trepte.2017d, the need for privacy is a secondary need---it is not an end in itself, but rather a way to satisfy other more fundamental needs (e.g., time to oneself). Specifically, @Westin.1967 defined four ultimate purposes of privacy: (1) self-development (i.e., the integration of experiences into meaningful patterns), (2) autonomy (i.e., the desire to avoid being manipulated and dominated), (3) emotional release (i.e., the release of tension from social role demands), and (4) protected communication (i.e., the ability to foster intimate relationships). Not least, privacy facilitates self-disclosure [@Dienlin.2014], which is vital for attaining social support, initiating relationships, and getting close to other people [@Omarzu.2000].

@@

# Method

@@

Summarizing across the results, the predictors explained `r inspect(results_threefactors_fit, 'r2')["pri_nee_gov"] * 100 %>% round(0)`% of the variance in need for privacy from the government, `r inspect(results_threefactors_fit, 'r2')["pri_nee_int"] * 100 %>% round(0)`% of the variance in need for privacy from other people, and `r inspect(results_threefactors_fit, 'r2')["pri_nee_ano"] * 100 %>% round(0)`% of the variance in need for anonymity. All predictors were also above our predefined SESOI of |$\upbeta$| = .10, supporting their theoretical relevance. In addition, note that several nonsignificant effects were actually larger than the predefined SESOI (e.g., the potential effect of risk avoidance on the need for privacy from the government). Hence, it is plausible that a larger study, which yields smaller standard errors, would find additional effects.

As stated above, because all predictors were included on the basis of specific a priori rationales we did not control for alpha error inflation. To better assess whether the results could also be explained by chance alone, let us briefly consider that if the null hypotheses had been true, given the predefined alpha level of 5 % and the `r nrow(t_results_threefactors)` statistical tests that were run, one would have expected to find `r round(nrow(t_results_threefactors) * .05, 2)` statistically significant results by chance alone. In this study we found `r round(nrow(filter(coeffs_tab(results_threefactors_fit), p < .05)), 0)` statistically significant results, which is substantially higher. Similarly, the average _p_-value _p_ = `r coeffs_tab(results_threefactors_fit)$p %>% mean %>% my_round("std")` was substantially smaller than the average _p_-value of _p_ = .50 that one would have expected to find if the null hypotheses had been true.

@@

Next, and more controversial, our results also showed that the desire for anonymity is related substantially with measures of integrity. People who self-reported lower integrity desired more anonymity. Respondents who said, for example, that they would feel tempted to take things that do not belong to them were also more likely to avoid situations in which they were identifiable. This finding hence supports the logic behind the “nothing to hide” argument. It also follows Altman’s privacy regulation theory [-@Altman.1976], which states that if exposure of information is risky it is likely that people will use more mechanisms to strengthen their social boundaries and increase their desired level of privacy.	Importantly, however, integrity was not related to need for privacy from other people or to need for privacy from government surveillance, which hence limits the generalizability of the nothing-to-hide argument. In addition, note that the responses to the items measuring need for anonymity were much lower compared to the other two dimensions of need for privacy (i.e., _M_~anonymity~ = `r t_psyc["Anonymity", "m"]`, _M_~government~ = `r t_psyc["Government", "m"]`, _M_~interpersonal~ = `r t_psyc["Interpersonal", "m"]`). Hence, the relationship between integrity and need for privacy might show only in very specific domains that are related to anonymity (for example, avoiding places with video surveillance). In conclusion, the relation between integrity and need for privacy seems to be limited to certain aspects of need for privacy only. It should thus not be overly generalized.

@@

Finally, in the SEM no significant relations between need for privacy and traditionality were found. The results imply that independent from being more or less open to change, people desire the same level of privacy. Taken further, this finding might illustrate that people who are more open to change, who are more likely to use novel devices, media, and services, or who might consider themselves as early adopters, do not need less privacy.

@@

Another limitation of the study is that despite using mostly well-established scales, confirmatory factor analyses revealed that most measures did not show satisfactory factorial validity, which resembles the current finding and criticism by @Hussey.2018. In order to achieve sufficient factorial validity we therefore needed to adapt the scales (for example by deleting items). However, this introduces problems of overfitting, which in turn potentially impedes the generalizability of the results [see, e.g., @Yarkoni.2017]. By using bifactor models we explicitly tried to retain the largest possible number of items, which should render our results more robust. In addition, also when looking at the items individually (see the OSM) one can see that the results are not overly dependent on inclusion of specific items.

Another methodological improvement for future studies concerns the sample, which was comparatively small, leading to a reduced power to detect also small effects. Nonetheless, we are still confident that the results should be reliable because we deliberately selected specific predictors that should exhibit at least small to moderate relations with need for privacy (i.e., _r_ = .20). For small to moderate relations, the current study still had a convincing power of `r power_smallmoderate`%. Granted, when aiming to find also small effects (i.e., _r_ = .10) with a probability of 95%, future studies would need to collect data from `r sample_95` respondents [@Cohen.1992]. Finally, generalizability of our sample is limited, as it was predominantly young, female, highly educated, and collected only in one country.
